Staged Abstract Interpreters

### 2019-09-11 paper/40_absinterp.tex
misc
polish

After seeing the unstaged and staged concrete interpreter, now we turn our
focus to abstract interpreters under the same framework. We first present a
lattice representation with binding-time abstraction, as well as simple
abstract domains, such as power sets.  Then, we show the abstract
components, namely, abstract values, abstract environments, and abstract stores.
The abstract interpreter we construct in this section is similar to
\citet{DBLP:journals/pacmpl/DaraisLNH17}'s. For simplicity, it is
context/path/flow-insensitive by our choice. In Section \ref{cfa}, we will
instantiate it to a commonly-used context-sensitive analysis;
\citet{Darais:2015:GTM:2814270.2814308} showed how to achieve path- and
flow-sensitivity by varying the monads, which is also applicable to our
unstaged or staged abstract interpreter.

@@

For function applications, @ap_clo@ looks the same as the concrete interpreter.
The difference is that now the first argument of @ap_clo@ is a set of closures
@funs@, so we can not directly destruct it to a syntactic @Lam@ by pattern
matching. Instead, we use an auxiliary function @lift_nd@ that takes a set and
lifts the elements in the set into the monad stack. Then we can
straightforwardly implement the @ap_clo@ in monadic style, where the closures
come from the monads and thus the nondeterminism can be naturally handled. The
light gray code shows what is added from the concrete @ap_clo@.
\begin{lstlisting}[escapechar=!]
  def lift_nd[T](vs: Set[T]): AnsM[T]
  def ap_clo(ev: Expr => Ans)(funs: Value, rand: Value): Ans = for {
    !\hl{CloV(Lam(x, e), $\rho$) <- lift\_nd[AbsValue](funs)}!
    α <- alloc(x)
    _ <- set_store(α → funs)
    rt <- local_env(ev(e))(ρ + (x → α))
  } yield rt
\end{lstlisting}

>>>>

After seeing the unstaged and staged concrete interpreter, now we turn our
focus to abstract interpreters under the same framework. We first present a
lattice representation with binding-time abstraction, as well as simple
abstract domains, such as power sets.  Then, we show the abstract
components, namely, abstract values, abstract environments, and abstract stores.
The abstract interpreter we construct in this section is similar to
\citet{DBLP:journals/pacmpl/DaraisLNH17}'s. For simplicity, it is
context/path/flow-insensitive by our choice. In Section \ref{cfa}, we will
instantiate it to a commonly-used context-sensitive analysis.
\citet{Darais:2015:GTM:2814270.2814308} showed how to achieve path- and
flow-sensitivity by varying the monads, which is also applicable to our
unstaged or staged abstract interpreter.

@@

For function applications, @ap_clo@ looks the same as the concrete interpreter.
The difference is that now the first argument of @ap_clo@ is a set of closures
@funs@, so we can not directly destruct it to a syntactic @Lam@ by pattern
matching. Instead, we use an auxiliary function @lift_nd@ that takes a set and
lifts the elements in the set into the monad stack. Then we can
straightforwardly implement the @ap_clo@ in monadic style, where the closures
come from the monads and thus the nondeterminism can be naturally handled. The
light gray line shows what is added from the concrete @ap_clo@.
\begin{lstlisting}[escapechar=!]
  def lift_nd[T](vs: Set[T]): AnsM[T]
  def ap_clo(ev: Expr => Ans)(funs: Value, rand: Value): Ans = for {
    !\hl{CloV(Lam(x, e), $\rho$) <- lift\_nd[AbsValue](funs)}!
    α <- alloc(x)
    _ <- set_store(α → funs)
    rt <- local_env(ev(e))(ρ + (x → α))
  } yield rt
\end{lstlisting}
